Depending on the specific application, online traveling agencies may offer services which
include transportation and accommodation. The focus of this section is to create an extensive
overview of the search tools available for finding commercial flights. 
In order to classify the search tools, the results will be grouped 
according to the \textit{trip type}:

\begin{itemize}
  \item single/round flight;
  \item multi-city trip;
\end{itemize}

This overview will also focus on the different search utilities provided 
by the search applications. Thus, it will analyze the ability to:

\begin{itemize}
  \item search over a range of start dates - \textbf{DR};
  \item present a price overview as a function of time - \textbf{PO};
  \item present different results for different criteria - \textbf{VR};
  \item display an interactive map - \textbf{Map};
\end{itemize}

There are other search utilities, presented by some online search applications,
which will not be subject of study. These include:

\begin{itemize}
  \item price monitoring and alerting service;
  \item analysis of the probability of price fluctuations;
  \item information about alternative flights;
  \item automatic caching for future usage;
\end{itemize}

This analysis will be considered for the major flight search applications,
which include, for example, Google Flights, Momondo, Skyscanner, Kayak, and Kiwi.



\subsubsection*{Single/Round flights}


In order to evaluate the search utilities provided by every flight search application,
as well as the actual price displayed, the same search will be performed over 
all applications. In particular, this search was executed over a month before the 
actual start date, and corresponds to:

\begin{itemize}
\item \textbf{S} - single flight - from Lisbon, to Amsterdam, at 8/6/2018;
\item \textbf{R} - round flight - from Lisbon, to Amsterdam, at 8/6/2018, for a duration of 3 days;
\end{itemize}

Table \ref{tab:simple_flights_analysis} displays 
the results of the search performed over the different applications,
together with a checklist specifying which search utilities are provided.
It is also worth noting that the price displayed in this table might not correspond to
the actual final price. For example, Edreams is well known for charging extra fees only upon the ticket purchase. 

\begin{table}[h]
  \centering
  \caption{Comparison of the search results, and search utilities, of different online flight search applications, for single and round flights.}
  \label{tab:simple_flights_analysis}
  \begin{tabular}{|l|l|l|l|l|l|l|}
  \hline
  company             & DR                & PO         & VR         & Map        & S                          & R                           \\ \hline
  Google Flights      & $\checkmark$              & $\checkmark$ &            & $\checkmark$ & 65                         & 127                         \\ \hline
  Expedia             &                              &            &            &            & 81                         & 167                         \\ \hline
  Booking.com / Kayak &                              &            & $\checkmark$ & $\checkmark$ & 67                         & 126                         \\ \hline
  Momondo             &                              & $\checkmark$ & $\checkmark$ &            & 62                         & \cellcolor[HTML]{C0C0C0}102 \\ \hline
  TripAdvisor         &                              &            &            &            & \cellcolor[HTML]{C0C0C0}56 & 124                         \\ \hline
  cheapflights.com    &                              &            &            &            & 75                         & 150                         \\ \hline
  Adioso              & $\checkmark$        &            & $\checkmark$ &            & 69                         & 137                         \\ \hline
  kiwi                & $\checkmark$        & $\checkmark$ & $\checkmark$ & $\checkmark$ & 76                         & 224                         \\ \hline
  SkyScanner          & $\checkmark$              & $\checkmark$ &            &            & 65                         & 127                         \\ \hline
  Edreams             & $\checkmark$              &            & $\checkmark$ &            & \cellcolor[HTML]{C0C0C0}56 & 116                         \\ \hline
  \end{tabular}
\end{table}

From the analysis of table \ref{tab:simple_flights_analysis}, it is possible to conclude that, for single and round flights:

\begin{itemize}
  \item different applications provide significantly different results. Single and round flight prices vary up to 44\% and 120\% respectively;
  \item half of the applications do not provide date range support;
  \item half of the applications do not provide price overview as a function of time;
  \item most applications do not present an interactive map;
\end{itemize}




\subsubsection*{Multi-city trips}


Following the same methodology as in the previous section, the same multi-city trip search will be performed 
on the different applications, and the result is presented as \textbf{M} in table \ref{tab:multi_city_analysis}. This search corresponds to:

\begin{itemize}
  \item single flight from Lisbon to Amsterdam, at 8/6/2018, for a duration of 3 days;
  \item single flight from Amsterdam to Berlin, at 11/6/2018, for a duration of 3 days;
  \item single flight from Berlin to Lisbon, at 14/6/2018;
\end{itemize}

It is worth noting that all flight search applications which currently perform multi-city trips,
do it in the exact same way: a multi-city trip corresponds to a collection of single flights,
which must specify the origin and destination, together with a single start date.

This case corresponds to a very inefficient search for a variety of reasons.
First of all, it constrains the search to one particular route. Second, it does not allow 
any kind of flexibility around the start date or duration of stay associated to each city. 
And finally, changing one parameter of the search might affect the rest of the trip.

Once again, the results of the search, together with the search utilities, are presented in table \ref{tab:multi_city_analysis}.
Note that MTCS stands for the ability to search for multi-city trips, as those specified above.

\begin{table}[h]
  \centering
  \caption{Comparison of the search results, and search utilities, of different online flight search applications, for multi-city trips.}
  \label{tab:multi_city_analysis}
  \begin{tabular}{|l|l|l|l|l|l|l|}
  \hline
  company             & MTCS       & DR  & PO         & VR         & Map                      & M                           \\ \hline
  Google Flights      & $\checkmark$ &       &            &            &                          & 184 \\ \hline
  Expedia             & $\checkmark$ &        &            &            &                          & 420                         \\ \hline
  Booking.com / Kayak & $\checkmark$ &        &            &            &                          & 191                         \\ \hline
  Momondo             & $\checkmark$ &        &            & $\checkmark$ &                          & \cellcolor[HTML]{C0C0C0}178 \\ \hline
  TripAdvisor         & $\checkmark$ &        &            & $\checkmark$ &  & 238                         \\ \hline
  cheapflights.com    &            &        &            &            &                          &                             \\ \hline
  Adioso              &            &        &            &            &                          &                             \\ \hline
  Kiwi                & $\checkmark$ &        & $\checkmark$ &            &                          & 226                         \\ \hline
  SkyScanner          & $\checkmark$ &        &            & $\checkmark$ &                          & 238                         \\ \hline
  Edreams             & $\checkmark$ &        &            &            &  & 333                         \\ \hline
  \end{tabular}
\end{table}

From the analysis of table \ref{tab:simple_flights_analysis}, it is possible to conclude that, for multi-city trips:

\begin{itemize}
  \item different applications provide significantly different results, which vary up to 135\%;
  \item most applications provide (limited) multi-city trip search; 
  \item most applications do not present different results for different search criteria;
  \item no application offers the ability to perform multi-city trips on a range of start dates; 
  \item no applications displays an interactive map;
\end{itemize}